An investigation into using scagnostics to find interesting projections of high-dimensional data

### 2018-10-26 scagnostics.Rmd
Replacing sphere with spiral.

```{r datasets}
set.seed(1984)
sphere100 <- sphereData(6, 100) %>% scale() %>% as_tibble()
sphere1000 <- sphereData(6, 1000) %>% scale() %>% as_tibble()

@@

```{r dataPlotsV1V2, fig.width=6, fig.height=7, fig.cap="Projection of all data sets considered onto the first two parameters V1 and V2, i.e. showing a typical uninformed view for each dataset."}
pL <- list()
i <- 1
pLabels <- c("Sphere 100", "Sphere 1000", "Pipe 100", "Pipe 1000", "Sine 100", "Sine 1000")
for(ds in list(sphere100, sphere1000, pipe100, pipe1000, sin100, sin1000)){
  pC <- ggplot(ds, aes(V1, V2)) +
    geom_point() +
    ggtitle(pLabels[i]) + theme(aspect.ratio=1)
  pL[[i]] <- pC
  i <- i+1
}
grid.arrange(pL[[1]], pL[[2]], pL[[3]], pL[[4]], pL[[5]], pL[[6]],
             ncol=2)
```

```{r dataPlotsV5V6, fig.width=6, fig.height=7, fig.cap="Projection of all data sets considered onto the last two parameters V5 and V6, i.e. showing the special view built into the Pipe and Sine datasets."}
pL <- list()
i <- 1
pLabels <- c("Sphere 100", "Sphere 1000", "Pipe 100", "Pipe 1000", "Sine 100", "Sine 1000")
for(ds in list(sphere100, sphere1000, pipe100, pipe1000, sin100, sin1000)){
  pC <- ggplot(ds, aes(V5, V6)) +
    geom_point() +
    ggtitle(pLabels[i]) + theme(aspect.ratio=1)
  pL[[i]] <- pC
  i <- i+1
}
grid.arrange(pL[[1]], pL[[2]], pL[[3]], pL[[4]], pL[[5]], pL[[6]],
             ncol=2)
```

@@

```{r dataPlotsDensity, fig.width=6, fig.height=7, fig.cap="One dimensional density distribution for all parameters and all data sets considered. Apart from statistical fluctuations we see that for the Pipe datasets the parameters V5 and V6 are pushed towards more extreme values by vetoing points away from the circle outline, and for the Sine dataset the parameter V6 follows a very distinct distribution. Note also that the first five paramters of the Sine dataset appear to be drawn from different normal distributions, which is a result of the rescaling, shifting the modes to different values for each direction."}
pL <- list()
i <- 1
pLabels <- c("Sphere 100", "Sphere 1000", "Pipe 100", "Pipe 1000", "Sine 100", "Sine 1000")
for(ds in list(sphere100, sphere1000, pipe100, pipe1000, sin100, sin1000)){
  distData <- melt(ds)
  pC <- ggplot(distData, aes(value, color=variable)) +
    geom_density() +
    ggtitle(pLabels[i])
  pL[[i]] <- pC
  i <- i+1
}
grid.arrange(pL[[1]], pL[[2]], pL[[3]], pL[[4]], pL[[5]], pL[[6]],
             ncol=2)
```

@@

```{r indexTable, echo=FALSE}
i <- 1
pLabels <- c("Sphere 100", "Sphere 1000", "Pipe 100", "Pipe 1000", "Sine 100", "Sine 1000")
indexOverview <- tibble(convex=numeric(),
               skinny=numeric(),
               stringy=numeric(),
               dcor2d=numeric(),
               splines2d=numeric(),
               MIC=numeric(),
               TIC=numeric(),
               holes=numeric(),
               name=character(),
               vars=character())
dsL <- list(sphere100, sphere1000, pipe100, pipe1000, sin100, sin1000)
for(ds in dsL){
  for(vars in list(c("V1","V2"),c("V5","V6"),c("V1","V6"))){
    dprj <- as.matrix(select_(ds, vars[1], vars[2]))
    scagRes <- scagnostics(dprj)
    dcorRes <- dcor2d(dprj[,1], dprj[,2])
    splineRes <- splines2d(dprj[,1], dprj[,2])
    mineRes <- mine(dprj[,1], dprj[,2])
    holesRes <- holes(dprj)
    indexOverview <- add_row(indexOverview, convex=scagRes[,"Convex"],
                    skinny=scagRes[,"Skinny"], stringy=scagRes[,"Stringy"],
                    dcor2d=dcorRes, splines2d=splineRes,
                    MIC=mineRes$MIC, TIC=mineRes$TIC, holes=holesRes,
                    name=pLabels[i], vars=paste0(vars[1],"-",vars[2]))
  }
  i <- i+1
}
knitr::kable(indexOverview,  caption = "Summary of index values.", digits=2) 
```

@@

```{r V1V2toV3V4}
if(!file.exists("cache/V1V2toV3V4.rda")){
  m1 <- matrix(c(1,0,0,0,0,0,0,1,0,0,0,0),ncol=2)
  m2 <- matrix(c(0,0,1,0,0,0,0,0,0,1,0,0),ncol=2)
  #this is silly but seems that first two entries are being ignored, so need some fake entries
  m3 <- matrix(c(1,0,0,0,1,0,0,1,0,0,0,1),ncol=2)
  m4 <- matrix(c(1,1,1,0,0,0,0,1,1,0,0,0),ncol=2)
  t1 <- save_history(sin100,tour_path=planned_tour(list(m3,m4,m1,m2)))
  t1full <- as.list(interpolate(t1))
  fullRes <- getProj(sin100, t1full, "sin", 100) %>%
    rbind(getProj(sin1000, t1full, "sin", 1000)) %>%
    rbind(getProj(sphere100, t1full, "sphere", 100)) %>%
    rbind(getProj(sphere1000, t1full, "sphere", 1000)) %>%
    rbind(getProj(pipe100, t1full, "pipe", 100)) %>%
    rbind(getProj(pipe1000, t1full, "pipe", 1000))
  save(t1, t1full, fullRes, file = "cache/V1V2toV3V4.rda")
} else {
  load("cache/V1V2toV3V4.rda")
}
```

@@

```{r plannedtourpath}
if(!file.exists("cache/plannedtourpath.rda")){
  m1 <- matrix(c(1,0,0,0,0,0,0,1,0,0,0,0),ncol=2)
  m2 <- matrix(c(0,0,0,0,1,0,0,0,0,0,0,1),ncol=2)
  #this is silly but seems that first two entries are being ignored, so need some fake entries
  m3 <- matrix(c(1,0,0,0,1,0,0,1,0,0,0,1),ncol=2)
  m4 <- matrix(c(1,1,1,0,0,0,0,1,1,0,0,0),ncol=2)
  t2 <- save_history(sin100,tour_path=planned_tour(list(m3,m4,m1,m2)))
  t2full <- as.list(interpolate(t2))
  fullResPlanned <- getProj(sin100, t2full, "sin", 100) %>%
    rbind(getProj(sin1000, t2full, "sin", 1000)) %>%
    rbind(getProj(sphere100, t2full, "sphere", 100)) %>%
    rbind(getProj(sphere1000, t2full, "sphere", 1000)) %>%
    rbind(getProj(pipe100, t2full, "pipe", 100)) %>%
    rbind(getProj(pipe1000, t2full, "pipe", 1000))
  save(t2, t2full, fullResPlanned, file = "cache/plannedtourpath.rda")
} else {
  load("cache/plannedtourpath.rda")
}
```

@@

```{r timer}
if(!file.exists("cache/timer.rda")){
  set.seed(2018)
  grandTour100 <- save_history(sphere100, grand_tour(2), max=4) %>%
    interpolate() %>%
    as.list()

>>>>

```{r datasets}
set.seed(1984)
spiral100 <- spiralData(6, 100) %>% scale() %>% as_tibble()
spiral1000 <- spiralData(6, 1000) %>% scale() %>% as_tibble()
#sphere100 <- sphereData(6, 100) %>% scale() %>% as_tibble()
#sphere1000 <- sphereData(6, 1000) %>% scale() %>% as_tibble()

@@

```{r dataPlotsV1V2, fig.width=6, fig.height=7, fig.cap="Projection of all data sets considered onto the first two parameters V1 and V2, i.e. showing a typical uninformed view for each dataset."}
pL <- list()
i <- 1
pLabels <- c("Spiral 100", "Spiral 1000", "Pipe 100", "Pipe 1000", "Sine 100", "Sine 1000")
for(ds in list(spiral100, spiral1000, pipe100, pipe1000, sin100, sin1000)){
  pC <- ggplot(ds, aes(V1, V2)) +
    geom_point() +
    ggtitle(pLabels[i]) + theme(aspect.ratio=1)
  pL[[i]] <- pC
  i <- i+1
}
grid.arrange(pL[[1]], pL[[2]], pL[[3]], pL[[4]], pL[[5]], pL[[6]],
             ncol=2)
```

```{r dataPlotsV5V6, fig.width=6, fig.height=7, fig.cap="Projection of all data sets considered onto the last two parameters V5 and V6, i.e. showing the special view built into the Pipe and Sine datasets."}
pL <- list()
i <- 1
pLabels <- c("Spiral 100", "Spiral 1000", "Pipe 100", "Pipe 1000", "Sine 100", "Sine 1000")
for(ds in list(spiral100, spiral1000, pipe100, pipe1000, sin100, sin1000)){
  pC <- ggplot(ds, aes(V5, V6)) +
    geom_point() +
    ggtitle(pLabels[i]) + theme(aspect.ratio=1)
  pL[[i]] <- pC
  i <- i+1
}
grid.arrange(pL[[1]], pL[[2]], pL[[3]], pL[[4]], pL[[5]], pL[[6]],
             ncol=2)
```

@@

```{r dataPlotsDensity, fig.width=6, fig.height=7, fig.cap="One dimensional density distribution for all parameters and all data sets considered. Apart from statistical fluctuations we see that for the Pipe datasets the parameters V5 and V6 are pushed towards more extreme values by vetoing points away from the circle outline, and for the Sine dataset the parameter V6 follows a very distinct distribution. Note also that the first five paramters of the Sine dataset appear to be drawn from different normal distributions, which is a result of the rescaling, shifting the modes to different values for each direction."}
pL <- list()
i <- 1
pLabels <- c("Spiral 100", "Spiral 1000", "Pipe 100", "Pipe 1000", "Sine 100", "Sine 1000")
for(ds in list(spiral100, spiral1000, pipe100, pipe1000, sin100, sin1000)){
  distData <- melt(ds)
  pC <- ggplot(distData, aes(value, color=variable)) +
    geom_density() +
    ggtitle(pLabels[i])
  pL[[i]] <- pC
  i <- i+1
}
grid.arrange(pL[[1]], pL[[2]], pL[[3]], pL[[4]], pL[[5]], pL[[6]],
             ncol=2)
```

@@

```{r indexTable, echo=FALSE}
i <- 1
pLabels <- c("Spiral 100", "Spiral 1000", "Pipe 100", "Pipe 1000", "Sine 100", "Sine 1000")
indexOverview <- tibble(convex=numeric(),
               skinny=numeric(),
               stringy=numeric(),
               dcor2d=numeric(),
               splines2d=numeric(),
               MIC=numeric(),
               TIC=numeric(),
               holes=numeric(),
               name=character(),
               vars=character())
dsL <- list(spiral100, spiral1000, pipe100, pipe1000, sin100, sin1000)
for(ds in dsL){
  for(vars in list(c("V1","V2"),c("V5","V6"),c("V1","V6"))){
    dprj <- as.matrix(select_(ds, vars[1], vars[2]))
    scagRes <- scagnostics(dprj)
    dcorRes <- dcor2d(dprj[,1], dprj[,2])
    splineRes <- splines2d(dprj[,1], dprj[,2])
    mineRes <- mine(dprj[,1], dprj[,2])
    holesRes <- holes(dprj)
    indexOverview <- add_row(indexOverview, convex=scagRes[,"Convex"],
                    skinny=scagRes[,"Skinny"], stringy=scagRes[,"Stringy"],
                    dcor2d=dcorRes, splines2d=splineRes,
                    MIC=mineRes$MIC, TIC=mineRes$TIC, holes=holesRes,
                    name=pLabels[i], vars=paste0(vars[1],"-",vars[2]))
  }
  i <- i+1
}
knitr::kable(indexOverview,  caption = "Summary of index values.", digits=2) 
```

@@

```{r V1V2toV3V4}
if(!file.exists("cache/V1V2toV3V4.rda")){
  m1 <- matrix(c(1,0,0,0,0,0,0,1,0,0,0,0),ncol=2)
  m2 <- matrix(c(0,0,1,0,0,0,0,0,0,1,0,0),ncol=2)
  #this is silly but seems that first two entries are being ignored, so need some fake entries
  m3 <- matrix(c(1,0,0,0,1,0,0,1,0,0,0,1),ncol=2)
  m4 <- matrix(c(1,1,1,0,0,0,0,1,1,0,0,0),ncol=2)
  t1 <- save_history(sin100,tour_path=planned_tour(list(m3,m4,m1,m2)))
  t1full <- as.list(interpolate(t1))
  fullRes <- getProj(sin100, t1full, "sin", 100) %>%
    rbind(getProj(sin1000, t1full, "sin", 1000)) %>%
    rbind(getProj(spiral100, t1full, "spiral", 100)) %>%
    rbind(getProj(spiral1000, t1full, "spiral", 1000)) %>%
    rbind(getProj(pipe100, t1full, "pipe", 100)) %>%
    rbind(getProj(pipe1000, t1full, "pipe", 1000))
  save(t1, t1full, fullRes, file = "cache/V1V2toV3V4.rda")
} else {
  load("cache/V1V2toV3V4.rda")
}
```

@@

```{r plannedtourpath}
if(!file.exists("cache/plannedtourpath.rda")){
  m1 <- matrix(c(1,0,0,0,0,0,0,1,0,0,0,0),ncol=2)
  m2 <- matrix(c(0,0,0,0,1,0,0,0,0,0,0,1),ncol=2)
  #this is silly but seems that first two entries are being ignored, so need some fake entries
  m3 <- matrix(c(1,0,0,0,1,0,0,1,0,0,0,1),ncol=2)
  m4 <- matrix(c(1,1,1,0,0,0,0,1,1,0,0,0),ncol=2)
  t2 <- save_history(sin100,tour_path=planned_tour(list(m3,m4,m1,m2)))
  t2full <- as.list(interpolate(t2))
  fullResPlanned <- getProj(sin100, t2full, "sin", 100) %>%
    rbind(getProj(sin1000, t2full, "sin", 1000)) %>%
    rbind(getProj(spiral100, t2full, "spiral", 100)) %>%
    rbind(getProj(spiral1000, t2full, "spiral", 1000)) %>%
    rbind(getProj(pipe100, t2full, "pipe", 100)) %>%
    rbind(getProj(pipe1000, t2full, "pipe", 1000))
  save(t2, t2full, fullResPlanned, file = "cache/plannedtourpath.rda")
} else {
  load("cache/plannedtourpath.rda")
}
```

@@

```{r timer}
if(!file.exists("cache/timer.rda")){
  set.seed(2018)
  grandTour100 <- save_history(sphereData(6,100), grand_tour(2), max=4) %>%
    interpolate() %>%
    as.list()
